---
title: "Data Visualization"
output:
  html_document: default
  highlight: tango
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

Reference:

1. R markdown https://rmarkdown.rstudio.com/authoring_quick_tour.html
2. Visualization Tutorial:  https://r4ds.had.co.nz/data-visualisation.html
3. Preview: `Control + Shift + K`

```{r}
# install a package once, reload it every time you start a new session.
# library(tidyverse) 
```


```{r}
?mpg
# mpg$displ means a car's engine size (in litres)
# mpg$hwy means a car's fuel efficiency on the higway (measured in miles per gallon (mpg)). The lower, the worser.
```

ggplot() creates a coordinate system that you can add layers to. The first argument of ggplot() is the dataset to use in the graph. 

```{r}
# ggplot(data = mpg) creates an empty graph
ggplot(data = mpg)
```

complete your graph by adding (+) one or more layers to ggplot(). The function geom_point() adds a layer of points to your plot, i.e. geom_point() creates a scatterplot. Each geom function in ggplot2 takes a mapping argument, which defines how variables in your dataset are mapped to visual properties. 

The mapping argument is always paired with aes(), and the x and y arguments of aes() specify which variables to map to the x and y axes

```{r}
ggplot(data = mpg) + geom_point(mapping = aes(x = displ, y = hwy))
```

you can add a categorical variable to a scatterplot by mapping it to an aesthetic -- a visual property of the objects in your plot. And ggplot2 will 
- 1. automatically assign a unique level of the aesthetic 
     (here a unique color) to each unique value of the variable,a process known as scaling. 
- 2. add a legend that explains which values correspond to which classes.

1. By Color
e.g.you can map the colors of your points to the class variable to reveal the class of each car
```{r}
ggplot(data = mpg) + geom_point(mapping = aes(x = displ, y = hwy, color = class))
```

2. By Shape
ggplot2 will only use six shapes at a time. By default, additional groups will go unplotted when you use the shape aesthetic.
```{r}
ggplot(data = mpg) + geom_point(mapping = aes(x = displ, y = hwy, shape = class)) 
```
We get a warning here, because mapping an unordered variable (class) to an ordered aesthetic (size) is not a good idea.

3. By Size of Points
```{r}
ggplot(data = mpg) + geom_point(mapping = aes(x = displ, y = hwy, size = class)) 
```

4. By Opacity
```{r}
ggplot(data = mpg) + geom_point(mapping = aes(x = displ, y = hwy, alpha = class))
```

5. Customized
You can also set the aesthetic properties of your geom manually by setting the aesthetic by name as an argument of your geom function; i.e. it goes **OUTSIDE** of aes()

```{r}
ggplot(data = mpg) + geom_point(mapping = aes(x = displ, y = hwy), color = "blue")
```

## Exercies

Map a continuous variable to color, size, and shape. How do these aesthetics behave differently for categorical vs. continuous variables?
```{r}
ggplot(data = mpg) + geom_point(mapping = aes(x = displ, y = hwy, color = cty))

ggplot(data = mpg) + geom_point(mapping = aes(x = displ, y = hwy, size = cty))

# Error: A continuous variable can not be mapped to shape
# ggplot(data = mpg) + geom_point(mapping = aes(x = displ, y = hwy, shape = cty))
```

What happens if you map the same variable to multiple aesthetics?
```{r}
ggplot(data = mpg) + geom_point(mapping = aes(x = displ, y = hwy, color = class, alpha = class))
```

What does the stroke aesthetic do? What shapes does it work with? (Hint: use ?geom_point)
```{r}
?geom_point

ggplot(data = mpg) + geom_point(mapping = aes(x = displ, y = hwy, stroke = cty))
```

What happens if you map an aesthetic to something other than a variable name, like aes(colour = displ < 5)? Note, you’ll also need to specify x and y.
```{r}
ggplot(data = mpg) + geom_point(mapping = aes(x = displ, y = hwy, color = displ < 5))
```

```{r}
ggplot(data = mpg) + geom_point(mapping = aes(x = displ, y = hwy)) + facet_wrap(~ class, nrow = 2)
ggplot(data = mpg) + geom_point(mapping = aes(x = displ, y = hwy)) + facet_grid(drv ~ cyl)
```