---
title: "R_tutorial"
output:
  pdf_document: default
  html_document: default
  highlight: tango_dark
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## R Markdown

Reference: 
https://rmarkdown.rstudio.com/authoring_quick_tour.html
https://bookdown.org/yihui/rmarkdown/pdf-document.html

Creating documents with R Markdown starts with an .Rmd file that contains a combination of markdown (content with simple text formatting) and R code chunks. The .Rmd file is fed to knitr ⧉, which executes all of the R code chunks and creates a new markdown (.md) document which includes the R code and it’s output.

When you click the **Knit** button a document will be generated that includes both content as well as the output of any embedded R code chunks within the document. You can embed an R code chunk like this:


```{r cars}
summary(cars)
```

## Including Plots

You can also embed plots, for example:

```{r pressure, echo=FALSE}
plot(pressure)
```

Note that the `echo = FALSE` parameter was added to the code chunk to prevent printing of the R code that generated the plot.
